---
title: "EDUC 641 Lab 1"
subtitle: "(Make friends with RStudio)"
author: "Congli (Claire) Zhang"
institute: "University of Oregon"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent()
```

# Housekeeping

### We are here

 - to support you with hands-on R programming tasks from lecture, textbook, and assignments
 - to give you a brief review of the code required for the current week
   
### We are NOT here 

 - to introduce new content or give you more tasks
 - to require your attendance or give you attendance points
 
---
# Housekeeping (cont'd)

### Two best practice of you 

 - come here with questions 
   - again, we don't have new content, your questions are best content for you
 
 - make friends with R. R is intimidating in the beginning, but over time, you'll see what a good friend it is in your journey of professional growth. Just like making friends:
   - make time for it every day
   - don't let critical misunderstandings stand in between you two
   - be patient
   - don't try to internalize everything about it
   - ...

---
# Today's checklist

1. (Pre-class), finish [EDUC R Workshop](https://community.uoregon.edu/courses/18778) modules 1 and 2 
2. Download and install R 
3. Download and install RStudio IDE
(Friendly note for folks who haven't done the above steps: a quick demo can be found in the [R Tutorial](https://daviddliebowitz.github.io/EDUC641_22F/slides/tutorial.html#5) slides on our corse website)
4. Open your RStudio IDE and familiarize yourself with its interface 

 - find the four panes 
     - code editor
     - R console
     - workspace and history
     - files and plots viewer
 - ask for help information 
     - go to your 4th pane "help" tab, click home button, see for yourself
     - ask specific information (what you've learned from R Workshop), e.g., ask for help information on basic R "mean" function

---
# Today's add-on

Hands-on activity in page #7 of the [R Tutorial](https://daviddliebowitz.github.io/EDUC641_22F/slides/tutorial.html#7)

  - and let us know if you have any questions


#### Feel free to leave after all activities and questions are cleared.
#### Thank you for attending this lab!





