---
title: "EDUC 641 Lab 3"
subtitle: "(Describing categorical data and exporting results)"
author: "Congli (Claire) Zhang"
institute: "University of Oregon"
date: "10/10 or 10/11/2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
library(countdown)
style_duo_accent()
```

### Today's objectives

1. Check your learning in R:

  - describing a categorical variable (present in a table)
  - visualizing a categorical variable (present in a plot)
  - interpret and summarize information from table and plot 
  
2. Learn how to save tables and plots out of RStudio

  - besides the "copy-paste" or "screenshot" strategy (okay to use for assignment 1 but let's get fancy today)

---
### First thing first

One of the best ways to learn in this class: did you successfully replicate Dr. Liebowitz's code from class OR at least run it on your own computer?

  - Where they are? For class #3, courseweb -> schedule -> unit1 -> lectures -> R code (download it NOW if haven't)
  
  - What you do? Open the script, run the code line-by-line, and see the results for yourself (we'll do this together in a sec)
  
  - What if? Highly likely you will:
     - find it hard to understand all the code - don't dwell on it. You'll revisit it again and again and good news, you'll know more about it each time.
     - run into some issues - best learning opportunity. Solving issues is the way you get better with R.
     
---
### Make the lecture code your own (practicing workflow from lab 2)

1. Create a new R project, call it **"class03"**
2. Create four folders, **"data", "code", "table", "plot"**
3. Go to courseweb, download the dataset **"deathpenalty.csv"** (where? courseweb -> schedule -> unit1 -> lectures -> data)
4. Put the dataset in your **data folder**
5. Within your R project, create a new R script (the file with .R extension), call it **"class3_code"**, and save it to your **code folder**
6. DIFFERENT FROM LAB 2, today we'll load a package in the beginning of your R script:
    - library(tidyverse)
7. Write code to load the dataset to your R script, call the dataframe **"df"** (to be consistent with lecture code - IMPORTANT)
8. You are now ready to copy (line #30 forward) from lecture code to your own script

`r countdown(minutes = 6, seconds = 0, left = 0, right = 0, padding = "5px", margin = "2%", font_size = "3em")`
---
### Major learning points from lecture code

1. Present a categorical variable in a table: run **code line #56 and #61** in your own R script
2. Visualize a categorical variable in a barplot in two ways:
    - Using base R (build-in functions, don't need any package): run **code line #57 and #62** in your own R script 
    - Using {ggplot2} package: run **code line #69 - #79** in your own R script 
3. Go around your table, answer questions below:
    - Describe the structure of the data (how many variables are there, what is the current type of each variable and what the type should be, how many rows/observations, etc.)
    - How many defendants were sentenced to death? In which table did you find this information? In which plot did you find this information?
    - What was the proportion of black victims? In which table did you find this information? In which plot did you find this information?
    
`r countdown(minutes = 6, seconds = 0, left = 0, right = 0, padding = "5px", margin = "2%", font_size = "3em")`
---
### Save your tables out of RStudio
(because most of the time, you need to share your tables and plots without the help of RStudio)

1. Find in the environment (or your previous code) the assigned names of the deathpenalty count table (hint: **counts**) and race proportion table (**prop**)
2. Write the below code to export the tables to .csv files and store them into your **table folder**
    - write.csv(counts, file = "./table/counts.csv", row.names = FALSE)
    - write.csv(prop, file = "./table/prop.csv", row.names = FALSE)
3. Go to your **table folder** and see if they are there

`r countdown(minutes = 6, seconds = 0, left = 0, right = 0, padding = "5px", margin = "2%", font_size = "3em")`

---
### Save your plots out of RStudio

1. Find in the environment (or your previous code) the assigned name of the plot (e.g., **rd**) 
2. Save all types of plot in the most intuitive way (**demo on my computer**) - which works nice for Base R plots
3. Save the plot created by {ggplot2}: write the below code to export the ggplot to .jpeg file and store it into your **plot folder**
    - ggsave(plot = rd, "./plot/race.jpeg")
4. Go to your **plot folder** and see if it's there


**Thank you for coming! Feel free to leave if finished.**



