---
title: "EDUC 610 Lab 5"
subtitle: "Describing and Visualizing Continuous Variables"
author: "Havi and Claire<br><span style = 'font-size: 50%;'>University of Oregon<br>Fall 2022</span>"
date: "10/24 or 10/25/2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(countdown)

style_duo_accent(
  primary_color = "#1F4E79",
  secondary_color = "#0072B2",
  header_font_google = xaringanthemer::google_font("Atkinson Hyperlegible", "600"),
  text_font_google   = xaringanthemer::google_font("Atkinson Hyperlegible", "200", "200i"),
  code_font_google   = xaringanthemer::google_font("IBM Plex Mono"),
  colors = c(
    darkblue = "#0072B2",
    lightblue = "#56B4E9",
    darkorange = "#D55E00",
    lightorange = "#E69f00",
    yellow = "#F0E442",
    green = "#009E73",
    pink = "#CC79A7",
    gray = "999999",
    white = "#FFFFFF"
  )
)

xaringanExtra::use_clipboard()
```

### Learning objectives

- R Skill

  - Data cleaning with `dplyr` verbs
      - `select()`
      - `filter()`
      - `mutate()`
      - Learn more by clicking [here](https://github.com/rstudio/cheatsheets/blob/main/data-transformation.pdf) 
      
  
  
*Note that {dplyr} is built in {tidyverse} so you don't need to load it separately.*
---
### Learning objectives - cont.

- Statistical Skills

  - Visualize continuous variables
     - `geom_histogram()` and `hist()`
     - `geom_density()`
     - `geom_boxplot()` and `boxplot()`
     
  - Describe conrinuous variables
      - `mean()`
      - `median()`
      - `range()`
      - `quantile()`
      - `IQR()`

---
### Before lab activities

1. Find today's slides in the mailbox. We will use life_expectancy.csv data. You can either watch or follow along. Simply copy the code from the slides to your R script.

2. The last **20 minutes** of the lab will be used for: 

  - Q&A about assignment 3 
  - Q&A about class
  - Self-working time

3. You should have downloaded Unit 3 lecture code (courseweb -> schedule -> unit3 -> lectures -> R code). If not, do it AFTER the lab

---
#### Task 1: Set up your working environment in R

First repeat what we've been doing over the past three weeks: create R project, make four folders, add life_expectancy.csv in the data folder and start an R script. 


Then load the packages and read the file.

```{r load-package, warning=FALSE, message=FALSE}
library(tidyverse)
```


```{r read file, echo = FALSE}
life <- read.csv("E:/Dropbox (University of Oregon)/Work@UO/Terms/2022-23/Fall22/GE/Class code/life_expectancy.csv")
```

```{r, eval = FALSE}
life <- read.csv("data/life_expectancy.csv")
```


---
#### Task 2: Let's inspect the structure

```{r}
dim(life)
colnames(life)
```

---
```{r}
str(life)
```

---
### `dplyr` verbs
#### `select()` the columns we want

Of the 22 variables, we are interested in the country, life_expectancy, status, and year columns. Let's keep only these.

General syntax - 

```{r, eval = FALSE}
new_name <- select(
  old_data, #name of dataset
  c(column names separated by comma) #names of columns as it appears 
)
```

---
### `dplyr` verbs
#### `select()` the columns we want

```{r}
life_short <- select(
  life, 
  c(Country, Year, Status, Life.expectancy)
)

dim(life_short)
```

You can also give index of the columns

```{r}
life_short <- select(
  life,
  c(1:4)
)

dim(life_short)
```

---
### `dplyr` verbs
#### `filter()` the rows using columns

**Brief detour**

Let's look at conditionals

```{r}
a <- 1
b <- 3
c <- 10
```


```{r, eval = FALSE}
a > b
a < b
a == b
```

---

```{r}
a > b
a < b
a == b
```

---

```{r}
a <- 1
b <- 3
c <- 10
```

```{r, eval = FALSE}
# | is the or operator
(a > b) | (c > b)
# read as a greater than b or c greater than b

# & is the and operator
(a > b) & (c > b)
# read as a greater than b and c greater than b
```

---


```{r}
#| is the or operator
(a > b) | (c > b)
# read as a greater than b or c greater than b

#& is the and operator
(a > b) & (c > b)
# read as a greater than b and c greater than b
```

---
### `dplyr` verbs
#### `filter()` the rows using columns - cont.

General syntax -

```{r, eval = FALSE}
new_name <- filter(
  old_name,
  (some condition on the columns)
)
```

---
### `dplyr` verbs
#### `filter()` the rows using columns - cont.

Say, we only want to look at developed countries.

```{r}
unique(life_short$Status)
```

```{r}
life_developed <- filter(
  life,
  Status == "Developed"
)

dim(life_developed)
```
---

### Combining `filter()` and `select()`

Say, we want to look at the these 4 columns- country, life_expectancy, status, and year columns for developed countries only.


#### Approach 1
```{r}
life_short <- select(
  life, 
  c(Country, Year, Status, Life.expectancy)
)

dim(life_short)

life_developed <- filter(
  life_short,
  Status == "Developed"
)

dim(life_developed)
```

---
### Combining `filter()` and `select()`

Say, we want to look at the these 4 columns- country, life_expectancy, status, and year columns for developed countries only.


#### Approach 2 - Using pipe %>% 

**First apply the first function. Then, apply the second function on the output of the first function** 
 
```{r}
life_short_developed <- life %>% 
  select(c(Country, Year, Status, Life.expectancy)) %>% 
  filter(Status == "Developed")

dim(life_short_developed)
```
**First select the four columns of life. Then, filter those columns where Status is developed **
---
### `dplyr` verbs
#### `mutate()` to the data frame


Basically, add a new column to the data frame based on existing values
Or, modify one of the existing values. 

General syntax - 
```{r, eval = FALSE}
new_name <- mutate(
  old_name,
  new_column_name = (some operation on the exisiting columns)
)
```

Or,

```{r, eval = FALSE}
new_name <- old_name %>% 
  mutate(
  new_column_name = (some operation on the exisiting columns)
)
```

---
### `dplyr` verbs
#### `mutate()` to the data frame

```{r}
life_short_developed$Life.expectancy[1:10]
```

Say, we want to round Life.expectancy to nearest integer

```{r}
life_short_developed <- life_short_developed %>% 
  mutate(Life.expectancy = round(Life.expectancy))

life_short_developed$Life.expectancy[1:10]
```

*Note: This is not a very useful example. `mutate()`. In this week's lectures and next week's lab, we will show more useful purposes of `mutate()`.*
---
### Let's visualize Life.expectancy

General syntax if using `ggplot2()`

```{r, eval = FALSE}
ggplot(data_file, aes(x = variable_name)) +
  geom_type()
```

General syntax if using `base` functions

```{r, eval = FALSE}
function_name(data_name$column_name)
```
---
#### Histogram

```{r,  warning=FALSE, message = FALSE, fig.height=6}
ggplot(life_short_developed, aes(x = Life.expectancy)) +
  geom_histogram()
```

---

#### Boxplot

```{r, warning=FALSE, message = FALSE, fig.height=6}
ggplot(life_short_developed, aes(y = Life.expectancy)) +
  geom_boxplot()
```

---
#### Density plot

```{r, warning=FALSE, message = FALSE, fig.height=6}
ggplot(life_short_developed, aes(x = Life.expectancy)) +
  geom_density()
```
---
### baseR alternative

#### Histogram
```{r, warning=FALSE, message = FALSE, fig.height=6}
hist(life_short_developed$Life.expectancy)
```

#### Boxplot
---
```{r, warning=FALSE, message = FALSE, fig.height=6}
boxplot(life_short_developed$Life.expectancy)
```

---
#### Let's make our histogram prettier

```{r, warning=FALSE, message = FALSE, fig.height=6, eval = FALSE}
ggplot(life_short_developed, aes(x = Life.expectancy)) +
  geom_histogram(fill = "cornflowerblue",
                 bins = 15)+
  labs(
    x = "Life expectancy",
    y = "Frequency",
    title = "Distribution of Life expectancy"
    # can also give subtitle and caption
  )+
  theme_minimal()
```
---

#### Let's make our histogram prettier

```{r, warning=FALSE, message = FALSE, fig.height=6, echo = FALSE}
ggplot(life_short_developed, aes(x = Life.expectancy)) +
  geom_histogram(fill = "cornflowerblue",
                 bins = 15)+
  labs(
    x = "Life expectancy",
    y = "Frequency",
    title = "Distribution of Life expectancy"
    # can also give subtitle and caption
  )+
  theme_minimal()
```
---
#### For curiosity

Wondering how many geoms? Check [here](https://ggplot2.tidyverse.org/reference/)

---
#### Some descriptives of the Life.expectancy variable

```{r}
mean(life_short_developed$Life.expectancy, na.rm = TRUE)
median(life_short_developed$Life.expectancy, na.rm = TRUE)
min(life_short_developed$Life.expectancy, na.rm = TRUE)
max(life_short_developed$Life.expectancy, na.rm = TRUE)
```

---
```{r}
range(life_short_developed$Life.expectancy,  na.rm = TRUE)
IQR(life_short_developed$Life.expectancy,  na.rm = TRUE)
quantile(life_short_developed$Life.expectancy,  na.rm = TRUE)
```

---
#### (If time allows) CHALLENGE

**Visualize and describe the Adult.Mortality variable**

  - Select the relevant columns and save it as `adult` object.
  - Filter the dataset include data from 2001-2010
  - Show the distribution of Adult.Mortality using any one geom
  - Use one summary statistic to describe the variable
  
---
#### One way to do the challenge

```{r}
adult <- life %>% 
  select(c(Country, Year, Status, Adult.Mortality)) %>% 
  filter(Year >= 2001 & Year <= 2010)

dim(adult)
head(adult)
```

---
#### One way to do the challenge

```{r, eval = FALSE}
ggplot(adult, aes(y = Adult.Mortality)) +
  geom_boxplot(
    fill = "cornflowerblue",
    color = "gray70",
    width = 0.25,
    outlier.colour = "red"
  ) +
  xlim(-0.4, 0.4)+
  labs(
    x = "",
    y = "Adult Mortality Rate per 1,000",
    title = "Distribution of Adult Mortality",
    subtitle = "The boxplot displays the distribution for Developed countries between 2000-2010"
  )+
  theme_classic()+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
  

```
---
#### One way to do the challenge

```{r, echo = FALSE}
ggplot(adult, aes(y = Adult.Mortality)) +
  geom_boxplot(
    fill = "cornflowerblue",
    color = "gray70",
    width = 0.25,
    outlier.colour = "red"
  ) +
  xlim(-0.4, 0.4)+
  labs(
    x = "",
    y = "Adult Mortality Rate per 1,000",
    title = "Distribution of Adult Mortality",
    subtitle = "The boxplot displays the distribution for Developed countries between 2000-2010"
  )+
  theme_classic()+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
  

```

---
#### One way to do the challenge

```{r}
mean(adult$Adult.Mortality)
```

