---
title: "EDUC 610 Lab 5"
subtitle: "Describing and Visualizing Continuous Variables"
author: "Havi Khurana<br><span style = 'font-size: 50%;'>University of Oregon<br>Fall 2022</span>"
date: "10/24 or 10/25/2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(countdown)

style_duo_accent(
  primary_color = "#1F4E79",
  secondary_color = "#0072B2",
  header_font_google = xaringanthemer::google_font("Atkinson Hyperlegible", "600"),
  text_font_google   = xaringanthemer::google_font("Atkinson Hyperlegible", "300", "300i"),
  code_font_google   = xaringanthemer::google_font("IBM Plex Mono"),
  colors = c(
    darkblue = "#0072B2",
    lightblue = "#56B4E9",
    darkorange = "#D55E00",
    lightorange = "#E69f00",
    yellow = "#F0E442",
    green = "#009E73",
    pink = "#CC79A7",
    gray = "999999",
    white = "#FFFFFF"
  )
)

xaringanExtra::use_clipboard()
```

### Learning objectives

- R Skill

  - Data cleaning with `dplyr` verbs
      - `select()`
      - `filter()`
      - `mutate()`
      - And, [many more](https://github.com/rstudio/cheatsheets/blob/main/data-transformation.pdf) 
      
  
*Note that {dplyr} is built in {tidyverse} so you don't need to load it separately.
---
### Learning objectives - cont.

- Statistical Skills

  - Visualize continuous variables
     - `geom_histogram()` and `hist()`
     - `geom_density()`
     - `geom_boxplot()` and `boxplot()`
     
  - Describe conrinuous variables
      - `mean()`
      - `median()`
      - `range()`
      - `quantile()`
      - `IQR()`

---
### Before lab activities

1. Find today's slides in the mailbox. We will use life_expectancy.csv data. You can either watch or follow along. Simply copy the code from the slides to your R script.

2. The last **20 minutes** of the lab will be used for: 

  - Q&A about assignment 3 
  - Q&A about class
  - Self-working time

3. You should have downloaded Unit 3 lecture code (courseweb -> schedule -> unit3 -> lectures -> R code). If not, do it AFTER the lab

---
#### Task 1: Set up your working environment in R
First repeat what we've been doing over the past three weeks: create R project, make four folders, add life_expectancy.csv in the data folder and start an R script. 


Then load the packages and read the file.

```{r load-package, warning=FALSE, message=FALSE}
library(tidyverse)
```


```{r read file, echo = FALSE}
life <- read.csv("E:/Dropbox (University of Oregon)/Work@UO/Terms/2022-23/Fall22/GE/Class code/life_expectancy.csv")
```

```{r, eval = FALSE}
life <- read.csv("data/life_expectancy.csv")
```


---
#### Task 2: Let's inspect the structure

```{r}
dim(life)
colnames(life)
```

---
```{r}
str(life)
```

---
### `dplyr` verbs
#### `select()` the columns we want

Of the 22 variables, we are interested in the country, life_expectancy, status, and year columns. Let's keep only these.

```{r}
life_short <- life %>% 
  #only column names without quotes
  select(Country, Year, Status, Life.expectancy)

head(life_short)
#Alternatively, give the index of column you want
life_short <- life %>% 
  select(c(1:4))
```

---
### `dplyr` verbs
#### `filter()` the rows using columns

**Brief detour**

Let's look at conditionals

```{r}
a <- 1
b <- 3
c <- 10
```


```{r, eval = FALSE}
a > b
a < b
a == b
```

---

```{r}
a > b
a < b
a == b
```

---

```{r}
a <- 1
b <- 3
c <- 10
```

```{r, eval = FALSE}
# | is the or operator
(a > b) | (c > b)
# read as a greater than b or c greater than b

# & is the and operator
(a > b) & (c > b)
# read as a greater than b and c greater than b
```

---


```{r}
#| is the or operator
(a > b) | (c > b)
# read as a greater than b or c greater than b

#& is the and operator
(a > b) & (c > b)
# read as a greater than b and c greater than b
```

---
### `dplyr` verbs
#### `filter()` the rows using columns - cont.

```{r}
unique(life_short$Status)
```

Say, we only want to look at developed countries.

```{r}
life_developed <- life_short %>% 
  filter(Status == "Developed") #Note the two equal signs. 
#And, the column name is not in quotes but the value is in quotes

head(life_developed)
```

---
### `dplyr` verbs
#### `filter()` the rows using columns - cont.

```{r}
unique(life_short$Year)
```

Say, we only want years between 2010 to 2015. How will we filter?

---
### `dplyr` verbs
#### `filter()` the rows using columns - cont.


```{r}
life_1015 <- life_short %>% 
  filter(Year >= 2010 & Year <= 2015)

head(life_1015)
```

---
### `dplyr` verbs
#### `mutate()` to the data frame


Basically, add a new column to the data frame based on existing values
Or, modify one of the existing values. 

```{r}
life_1015$Life.expectancy[1:10]
```
Say, we want to round Life.expectancy to neartest integer

```{r}
life_1015 <- life_1015 %>% 
  mutate(Life.expectancy = round(Life.expectancy))

life_1015$Life.expectancy[1:10]
```

---
### Let's visualize Life.expectancy

The base code remains the same if using `ggplot2()`

```{r, eval = FALSE}
ggplot(data_file, aes(x = variable_name)) +
  geom_type()
```

---
#### Histogram

```{r,  warning=FALSE, message = FALSE, fig.height=6}
ggplot(life_1015, aes(x = Life.expectancy)) +
  geom_histogram()
```

---

#### Boxplot

```{r, warning=FALSE, message = FALSE, fig.height=6}
ggplot(life_1015, aes(y = Life.expectancy)) +
  geom_boxplot()
```

---
#### Density plot

```{r, warning=FALSE, message = FALSE, fig.height=6}
ggplot(life_1015, aes(x = Life.expectancy)) +
  geom_density()
```
---
### baseR alternative

#### Histogram
```{r, warning=FALSE, message = FALSE, fig.height=6}
hist(life_1015$Life.expectancy)
```

#### Boxplot
---
```{r, warning=FALSE, message = FALSE, fig.height=6}
boxplot(life_1015$Life.expectancy)
```

---
#### Let's make our histogram prettier

```{r, warning=FALSE, message = FALSE, fig.height=6}
ggplot(life_1015, aes(x = Life.expectancy)) +
  geom_histogram(fill = "cornflowerblue")+
  labs(
    x = "Life expectancy",
    y = "Frequency",
    title = "Distribution of Life expectancy"
    # can also give subtitle and caption
  )+
  theme_minimal()
```
---
#### For curiosity

Wondering how many geoms? Check [here](https://ggplot2.tidyverse.org/reference/)

---
#### Some descriptives of the Life.expectancy variable

```{r}
mean(life_1015$Life.expectancy, na.rm = TRUE)
median(life_1015$Life.expectancy, na.rm = TRUE)
min(life_1015$Life.expectancy, na.rm = TRUE)
max(life_1015$Life.expectancy, na.rm = TRUE)
```

---
```{r}
range(life_1015$Life.expectancy,  na.rm = TRUE)
IQR(life_1015$Life.expectancy,  na.rm = TRUE)
quantile(life_1015$Life.expectancy,  na.rm = TRUE)
```

---
#### (If time allows) CHALLENGE

**Visualize and describe the Adult.Mortality variable**

  - First, select the relevant columns and save it as `adult` object.
  - Filter the dataset include data from 2001-2010
  - Show the distribution of the variable
  - Use one summary statistic to describe the variable
