---
title: "EDUC 610 Lab 6"
subtitle: "Describing and Visualizing Continuous Variables"
author: "Havi and Claire<br><span style = 'font-size: 50%;'>University of Oregon<br>Fall 2022</span>"
date: "10/31 or 11/01/2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(countdown)

style_duo_accent(
  primary_color = "#1F4E79",
  secondary_color = "#0072B2",
  header_font_google = xaringanthemer::google_font("Atkinson Hyperlegible", "600"),
  text_font_google   = xaringanthemer::google_font("Atkinson Hyperlegible", "200", "200i"),
  code_font_google   = xaringanthemer::google_font("IBM Plex Mono"),
  colors = c(
    darkblue = "#0072B2",
    lightblue = "#56B4E9",
    darkorange = "#D55E00",
    lightorange = "#E69f00",
    yellow = "#F0E442",
    green = "#009E73",
    pink = "#CC79A7",
    gray = "999999",
    white = "#FFFFFF"
  )
)

xaringanExtra::use_clipboard()
```

### Today's objectives

- Apply the concepts of last week's lab on Assignment 3

---

### Plan

- Self/group-working time
- I'll go around and help with questions/troubleshooting
- Refer to slides for hint/guidance 
- Caveat: The slides focus on the coding aspect of the assignment. In the assignment, you also have to interpret the output. Though this is not included in the slides, you're welcome to discuss it in your groups and with me. 

---

### Functions covered last week

- Data cleaning with `dplyr` verbs
    - `select()`
    - `filter()`
    - `mutate()` (lightly)

- Visualize continuous variables
    - `geom_histogram()` and `hist()`
    - `geom_density()`
    - `geom_boxplot()` and `boxplot()`
     
- Describe continuous variables
    - `mean()`
    - `median()`
    - `range()`
    - `quantile()`
    - `IQR()`

---

### Before beginning 

1. Find today's slides in the mailbox. The illustration uses the life_expectancy.csv data. The questions are parallel to Assignment 3, but applies on this dataset.

---

### Similar to Q 1.1

Open your RStudio, create a project and save it. Go to the root directory of the project and create folders named: “Code”, “Data”, “Figures” and “Tables.” Download the cont.csv dataset and store it in the folder “Data”. Create an R script (or .Rmd) file in the Code folder. Read the data into your R environment. You do not need to include this part of the response in your memo; only in your code.

You will need to load the packages and read the file.

```{r load-package, warning=FALSE, message=FALSE}
library(tidyverse)
```

```{r read file, echo = FALSE}
life <- read.csv("E:/Dropbox (University of Oregon)/2022-23/Fall22/GE/Class code/life_expectancy.csv")
```

*Instead of data use the name of the folder where you saved the file. Instead of life_expectancy, use the name you used to save the file.*

```{r, eval = FALSE}
life <- read.csv("data/life_expectancy.csv")
```


---

### Similar to Q 1.2 - Recode variables

Let's inspect these two continuous variables:
1. BMI
2. Schooling

*Create a smaller dataset `life1` the contains the following columns: Country, Year, BMI, Schooling. Change Country into type factor*

**HINT: Use the `select()` and `factor()` functions**

General syntax - 

```{r, eval = FALSE}
new_df <- select(
  old_df, #name of dataset
  c(column names/index separated by comma) 
  #names of columns as it appears and without quotes
)

#Or,
new_df <- old_df %>% 
    select(c(column names/index separated by comma))
```


---

### Similar to Q 1.2 - Inspect the data structure

Write your own code to view the structure of the dataset. How many rows/observations are there? How many variables are there? What type is each variable (after converting types)? What does each variable represent?

**HINT: There are many ways to do it. You can use the `str()`, `dim()`, `class()`, `typeof()`, `colnames()` functions**

---

### One way to do Q 1.1 and 1.2

```{r, eval = FALSE}
#Select relevant columns
life1 <- life %>% 
    select(c(Country, Year, BMI, Schooling))

#Check dimensions
dim(life1)

#Remove old df
rm(life)

#Change Country to factor
life1$Country <- factor(life1$Country)

#Note there are 193 countries, so I am letting R choose the factor labels and levels
#For the assignment, you will have to specify labels and levels

str(life1)
colnames(life1)
```


---


### Similar to Q2 - Descriptive Statistics

- Find the mean, median, range, iner-quartile range, variance and standard deviation of BMI and Schooling?
- Create a plot to show the distribution of BMI and Schooling. Describe the distribution, referencing your plot as appropriate. Make sure to use the following terms: tail(s), skew, kurtosis, mound/modal.

**HINT: There are several alternatives. You can use the `mean()`, `median()`, `IQR()`, `quantile()`, `range()`, `sd()`, `var()` functions. Remember to use `na.rm = TRUE` to remove missing values. You can draw histograms, boxplots, density plots, either using ggplot2 or base function.**

General syntax if using `ggplot2()`

```{r, eval = FALSE}
ggplot(data_file, aes(x = variable_name)) +
  geom_type()
```

General syntax if using `base` functions

```{r, eval = FALSE}
function_name(data_name$column_name)
```

---

### One way to do Q2

```{r, eval = FALSE}
#Descriptive Statistics of BMI
mean(life1$BMI, na.rm = TRUE)
median(life1$BMI, na.rm = TRUE)
sd(life1$BMI, na.rm = TRUE)
var(life1$BMI, na.rm = TRUE)
range(life1$BMI, na.rm = TRUE)
#diff(range(life1$BMI, na.rm = TRUE))
IQR(life1$BMI, na.rm = TRUE)
quantile(life1$BMI, na.rm = TRUE)

hist(life1$BMI)
ggplot(life1, aes(x = Schooling))+
    geom_density()+
    theme_bw()
```

---

### Similar to Q 3.1 - Transformations

Standardize the variables Schooling to a mean of 0 and standard deviation of 1.


---

#### Using $ sign

```{r, eval = FALSE}
Schooling.mean = mean(life1$Schooling, na.rm = TRUE)
Schooling.sd = sd(life1$Schooling, na.rm = TRUE)

life1$Schooling_zscore <- 
    (life1$Schooling - Schooling.mean)/Schooling.sd
```


---

#### Using `dplyr::mutate()` verbs

`dplyr::mutate()` add a new column to the data frame based on existing values

Or, modify one of the existing values. 

General syntax - 
```{r, eval = FALSE}
new_df <- mutate(
  old_df,
  new_column_name = (some operation on the exisiting columns)
)
```

Or,

```{r, eval = FALSE}
new_df <- old_df %>% 
  mutate(
  new_column_name = (some operation on the exisiting columns)
)
```

---

#### Using `dplyr::mutate()` verbs

```{r, eval = FALSE}
Schooling.mean = mean(life1$Schooling, na.rm = TRUE)
Schooling.sd = sd(life1$Schooling, na.rm = TRUE)

life1 <- life1 %>% 
  mutate(
      Schooling.std = (Schooling - Schooling.mean)/Schooling.sd
      )

ggplot(life1, aes(x = Schooling.std))+
    geom_density()+
    theme_bw()
```

 

---

### Similar to Q 3.2 - Examine one case

Let's examine one particular country, Australia, for the year 2010. Interpret the z-scores of Schooling and say how it compares in light of the distribution of the two variables.

**HINT: Use the `filter()` function**

General syntax - 

```{r, eval = FALSE}
new_df <- filter(
  old_df, #name of dataset
  some condition on the columns 
  #names of columns as it appears and without quotes
)

#Or,
new_df <- old_df %>% 
    filter(some condition on the columns)
```

---

### One way to do Q 3.2

```{r, eval = FALSE}
life1_aus <- life1 %>% 
    filter(Country == "Australia" & Year == "2011")

head(life1_aus)
```