---
title: "EDUC 610 Lab 6 & 7"
subtitle: "Describing and Visualizing Continuous Variables"
author: "Havi and Claire<br><span style = 'font-size: 50%;'>University of Oregon<br>Fall 2022</span>"
date: "10/31-11/8/22"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(countdown)

style_duo_accent(
  primary_color = "#1F4E79",
  secondary_color = "#0072B2",
  header_font_google = xaringanthemer::google_font("Atkinson Hyperlegible", "600"),
  text_font_google   = xaringanthemer::google_font("Atkinson Hyperlegible", "150", "150i"),
  code_font_google   = xaringanthemer::google_font("IBM Plex Mono"),
  colors = c(
    darkblue = "#0072B2",
    lightblue = "#56B4E9",
    darkorange = "#D55E00",
    lightorange = "#E69f00",
    yellow = "#F0E442",
    green = "#009E73",
    pink = "#CC79A7",
    gray = "999999",
    white = "#FFFFFF"
  )
)

xaringanExtra::use_clipboard()
```

### Today's objectives

- Personalized R support from Claire and Havi on Assignment 3
- Discuss interpretations in small groups

---

### Plan

- Self/group-working time
- We'll go around and help with questions/troubleshooting
- Refer to slides for hint/guidance 
---

### Functions covered in Week 5

- Data cleaning with `dplyr` verbs
    - `select()`
    - `filter()`
    - `mutate()` (lightly)

- Visualize continuous variables
    - `geom_histogram()` and `hist()`
    - `geom_density()`
    - `geom_boxplot()` and `boxplot()`
     
- Describe continuous variables
    - `mean()`
    - `median()`
    - `range()`
    - `quantile()`
    - `IQR()`

---

### Before beginning 

1. Find today's slides in the mailbox. The illustration uses the life_expectancy.csv data. The questions are parallel to Assignment 3, but applies on this dataset.

---

### Similar to Q 1.1

Open your RStudio, create a project and save it. Go to the root directory of the project and create folders named: “Code”, “Data”, “Figures” and “Tables.” Download the cont.csv dataset and store it in the folder “Data”. Create an R script (or .Rmd) file in the Code folder. Read the data into your R environment. You do not need to include this part of the response in your memo; only in your code.

You will need to load the packages and read the file.

```{r load-package, warning=FALSE, message=FALSE}
library(tidyverse)
```

```{r read file, echo = FALSE}
life <- read.csv("E:/Dropbox (University of Oregon)/2022-23/Fall22/GE/Class code/life_expectancy.csv")
```

*Instead of data use the name of the folder where you saved the file. Instead of life_expectancy, use the name you used to save the file.*

```{r, eval = FALSE}
life <- read.csv("data/life_expectancy.csv")
```


---

### Similar to Q 1.2
### Similar to Q 1.2 - Inspect the data structure

Write your own code to view the structure of the dataset. How many rows/observations are there? How many variables are there? What type is each variable (after converting types)? What does each variable represent?

**HINT: There are many ways to do it. You can use the `str()`, `dim()`, `class()`, `typeof()`, `colnames()` functions**

---

### One way to do Q 1.2

```{r, eval = FALSE}
#Check dimensions
dim(life)
#returns number of rows and columns in the dataset

#Change Country to factor with default levels
#Note there are 193 countries, so I am letting R choose the factor labels 
#and levels
life$Country <- factor(life$Country)

#Change Status to factors by specifying exact labels
#Factor variables with small number of levels can be coded manually
life$Status <- factor(life$Status,
                      levels = c("Developing", "Developed"),
                      labels = c("Developing", "Developed"))
```

---

### One way to do Q 1.2 

```{r, eval = FALSE}
#Structure of data.frame
str(life)
#Returns the column names and their type, as well as some values in that column

#Column names of data.frame
colnames(life)
#Returns the names of columns

#Storage type of vectors
typeof(life$BMI)
# typeof() returns one of integer, double, character, logical, list
# It basically tells how R is internally storing the information
# This is NOT what we are want when asking the type of variable in the assignments

#More informative description of the vector
class(life$BMI)
#class() returns a more useful description of the vector
#It can be numeric, integer, factor, array, matrix, table, data.frame, etc.
#It tells a bit more about the variable structure
```


---


### Similar to Q2 - Descriptive Statistics

- Find the mean, median, range, iner-quartile range, variance and standard deviation of BMI and Schooling?
- Create a plot to show the distribution of BMI and Schooling. Describe the distribution, referencing your plot as appropriate. Make sure to use the following terms: tail(s), skew, kurtosis, mound/modal.

**HINT: There are several alternatives. You can use the `mean()`, `median()`, `IQR()`, `quantile()`, `range()`, `sd()`, `var()` functions. Remember to use `na.rm = TRUE` to remove missing values. You can draw histograms, boxplots, density plots, either using ggplot2 or base function.**

General syntax if using `ggplot2()`

```{r, eval = FALSE}
ggplot(data_file, aes(x = variable_name)) +
  geom_type()
```

General syntax if using `base` functions

```{r, eval = FALSE}
function_name(data_name$column_name)
```

---

### One way to do Q2

```{r, eval = FALSE}
#Descriptive Statistics of BMI

#1. mean() returns the average of an numeric/integer column.
#na.rm is telling to remove the missing information in the column, 
#and then calculate the average
mean(life$BMI, na.rm = TRUE)

#2. median() returns the value at the 50th percentile of the data
median(life$BMI, na.rm = TRUE)

#3. sd() returns the standard deviation, a measure of spread of the distribution.
sd(life$BMI, na.rm = TRUE)

#4. var() returns the variance, another measure of the spread of a distribution. 
#It is basically the square of sd
var(life$BMI, na.rm = TRUE)
```

---

### One way to do Q2

```{r, eval = FALSE}
#5. range() returns the min and max observations, aka the upper and lower limit
#However, in statistics, range is max observation - min observation
#R doesn't return this number directly, you will have to subtract the 
#two values and find range yourself
range(life$BMI, na.rm = TRUE)

#6. quantile() returns the 0th, 25th, 50th, 75th, 100th percentile of a numeric column.
#Basically the minimum observation (0), 
#the observation that splits the distribution in 1 and 3 parts (25)
#The observation that splits in half (50), 
#the observation that splits in 3 and 1 parts (75), and the maximum (100)
#You can run this function to find the median, range, and inter-quartile range
quantile(life$BMI, na.rm = TRUE)

#7. IQR returns the difference of the 75th percentile and 25th percentile of a numeric column
#For the assignment, you have to find the upper and lower limits, as well as the range. 
#You should use quantile() to find the upper and lower limits

IQR(life$BMI, na.rm = TRUE)
```

---

### One way to do Q2

```{r, eval = FALSE}
#Plots using base and ggplot2

hist(life$BMI)
ggplot(life, aes(x = Schooling))+
    geom_density()+
    theme_bw()
```

---

### Similar to Q 3.1 - Transformations

Standardize the variables Schooling to a mean of 0 and standard deviation of 1.


---

#### Using $ sign

```{r, eval = FALSE}
#z-score is (observed value - mean)/(standard deviation)

#First, save the mean and sd in another variable to 
#avoid writing a big chunk of code together
Schooling.mean <- mean(life$Schooling, na.rm = TRUE)
Schooling.sd <- sd(life$Schooling, na.rm = TRUE)

#Now, use the formula of the z-score
life$Schooling_zscore <- 
    (life$Schooling - Schooling.mean)/Schooling.sd
```


---

#### Using `dplyr::mutate()` verbs

`dplyr::mutate()` add a new column to the data frame based on existing values

Or, modify one of the existing values. 

General syntax - 
```{r, eval = FALSE}
new_df <- mutate(
  old_df,
  new_column_name = (some operation on the exisiting columns)
)
```

Or,

```{r, eval = FALSE}
new_df <- old_df %>% 
  mutate(
  new_column_name = (some operation on the exisiting columns)
)
```

---

#### Using `dplyr::mutate()` verbs

```{r, eval = FALSE}
Schooling.mean <- mean(life$Schooling, na.rm = TRUE)
Schooling.sd <- sd(life$Schooling, na.rm = TRUE)

life <- life %>% 
  mutate(
      Schooling.std = (Schooling - Schooling.mean)/Schooling.sd
      )

#Plotting z-scores
ggplot(life, aes(x = Schooling.std))+
    geom_density()+
    theme_bw()
```

 

---

### Similar to Q 3.2 - Examine one case

Let's examine one particular country, Australia, for the year 2010. Interpret the z-scores of Schooling and say how it compares in light of the distribution of the two variables.

**HINT: Use the `filter()` function**

General syntax - 

```{r, eval = FALSE}
new_df <- filter(
  old_df, #name of dataset
  some condition on the columns 
  #names of columns as it appears and without quotes
)

#Or,
new_df <- old_df %>% 
    filter(some condition on the columns)
```

---

### One way to do Q 3.2

```{r, eval = FALSE}
life_aus <- life %>% 
    filter(Country == "Australia" & Year == "2011")

head(life_aus)
#z-scores tell you how many standard deviations away from the mean an observation is
#The further away the observed value is, the rarer it is to observe.
#This comes from the 68-95-99 rule
```

---
### Similar to Q 4

A report finds that the average global schooling is 8.7 years in 2022. We want to find if the average schooling in our same is different from 8.7.

```{r, eval =FALSE}
#we will use a one-sample t-test
#H0: the mean value of schooling is not different from 8.7 years.
#H1: the mean value of schooling is not 8.7 years

t.test(life$Schooling, mu = 8.7)
#You have to report the t-statistic, df, and p-values.
```

