---
title: "EDUC 610 Lab 6"
subtitle: "Describing and Visualizing Continuous Variables"
author: "Havi and Claire<br><span style = 'font-size: 50%;'>University of Oregon<br>Fall 2022</span>"
date: "10/31 or 11/01/2022"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
library(countdown)

style_duo_accent(
  primary_color = "#1F4E79",
  secondary_color = "#0072B2",
  header_font_google = xaringanthemer::google_font("Atkinson Hyperlegible", "600"),
  text_font_google   = xaringanthemer::google_font("Atkinson Hyperlegible", "200", "200i"),
  code_font_google   = xaringanthemer::google_font("IBM Plex Mono"),
  colors = c(
    darkblue = "#0072B2",
    lightblue = "#56B4E9",
    darkorange = "#D55E00",
    lightorange = "#E69f00",
    yellow = "#F0E442",
    green = "#009E73",
    pink = "#CC79A7",
    gray = "999999",
    white = "#FFFFFF"
  )
)

xaringanExtra::use_clipboard()
```

### Today's objectives

- Review functions we learnt last time
- Apply the learning on Assignment 3
- Go over `dplyr::mutate()`

---

### Plan

- Self/group-working time
- I'll go around and help with questions/troubleshooting
- Practice questions are aligned with Assignment 3 questions

---

### Last Week Review

- Data cleaning with `dplyr` verbs
    - `select()`
    - `filter()`
    - `mutate()` (lightly)

- Visualize continuous variables
    - `geom_histogram()` and `hist()`
    - `geom_density()`
    - `geom_boxplot()` and `boxplot()`
     
- Describe conrinuous variables
    - `mean()`
    - `median()`
    - `range()`
    - `quantile()`
    - `IQR()`

---

### Before lab activities

1. Find today's slides in the mailbox. We will use life_expectancy.csv data. The questions are similar to Assignment 3, but we will practice on a different data set.

2. You should have downloaded Unit 3 lecture code (courseweb -> schedule -> unit3 -> lectures -> R code). If not, do it AFTER the lab

---

### Task 1: Set up your working environment in R

First repeat what we've been doing over the past three weeks: create R project, make four folders, add life_expectancy.csv in the data folder and start an R script. 

Then load the packages and read the file.

```{r load-package, warning=FALSE, message=FALSE}
library(tidyverse)
```

```{r read file, echo = FALSE}
life <- read.csv("E:/Dropbox (University of Oregon)/2022-23/Fall22/GE/Class code/life_expectancy.csv")
```

*Instead of data use the name of the folder where you saved the file. Instead of life_expectancy, use the name you used to save the file.*

```{r, eval = FALSE}
life <- read.csv("data/life_expectancy.csv")
```


---

### Task 2: Subset the data and inspect structure

Let's inspect these two continuous variables:
1. BMI
2. Schooling

*Create a smaller dataset `life1` the contains the following columns: Country, Year, BMI, Schooling. Change Country into type factor*

**HINT: Use the `select()` and `factor()` functions**

General syntax - 

```{r, eval = FALSE}
new_df <- select(
  old_df, #name of dataset
  c(column names/index separated by comma) 
  #names of columns as it appears and without quotes
)

#Or,
new_df <- old_df %>% 
    select(c(column names/index separated by comma))
```


---

### Task 2: Inspect the data `life1`

Write your own code to view the structure of the dataset. How many rows/observations are there? How many variables are there? What type is each variable (after converting types)? What does each variable represent?

**HINT: There are many ways to do it. You can use the `str()`, `dim()`, `class()`, `typeof()`, `colnames()` functions**

---

### One way to do Task 2

```{r, eval = FALSE}
#Select relevant columns
life1 <- life %>% 
    select(c(Country, Year, BMI, Schooling))

#Check dimensions
dim(life1)

#Remove old df
rm(life)

#Change Country to factor
life1$Country <- factor(life1$Country)

str(life1)
colnames(life1)
```


---


### Task 3: Descriptive Statistic

- Find the mean, median, range, iner-quartile range, variance and standard deviation of BMI and Schooling?
- Create a plot to show the distribution of BMI and Schooling. Describe the distribution, referencing your plot as appropriate. Make sure to use the following terms: tail(s), skew, kurtosis, mound/modal.

**HINT: There are several alternatives. You can use the `mean()`, `median()`, `IQR()`, `quantile()`, `range()`, `sd()`, `var()` functions. Remember to use `na.rm = TRUE` to remove missing values. You can draw histograms, boxplots, density plots, either using ggplot2 or base function.**

General syntax if using `ggplot2()`

```{r, eval = FALSE}
ggplot(data_file, aes(x = variable_name)) +
  geom_type()
```

General syntax if using `base` functions

```{r, eval = FALSE}
function_name(data_name$column_name)
```

---

### One way to do Task 3

```{r, eval = FALSE}
#Descriptive Statistics of BMI
mean(life1$BMI, na.rm = TRUE)
median(life1$BMI, na.rm = TRUE)
sd(life1$BMI, na.rm = TRUE)
var(life1$BMI, na.rm = TRUE)
range(life1$BMI, na.rm = TRUE)
#diff(range(life1$BMI, na.rm = TRUE))
IQR(life1$BMI, na.rm = TRUE)
quantile(life1$BMI, na.rm = TRUE)

hist(life1$BMI)
ggplot(life1, aes(x = Schooling))+
    geom_density()+
    theme_bw()
```

---

### Task 4: Transformations

Standardize the variables Schooling a mean of 0 and standard deviation of 1.

---

### One way to do Task 4

```{r, eval = FALSE}
who$life_expectancy_zscore <- 
  (who$life_expectancy - mean(who$life_expectancy)) /
  sd(who$life_expectancy)
```


---

### `dplyr` verbs
### `mutate()` a new column to the data frame

Basically, add a new column to the data frame based on existing values
Or, modify one of the existing values. 

General syntax - 
```{r, eval = FALSE}
new_df <- mutate(
  old_df,
  new_column_name = (some operation on the exisiting columns)
)
```

Or,

```{r, eval = FALSE}
new_df <- old_df %>% 
  mutate(
  new_column_name = (some operation on the exisiting columns)
)
```

---

### Using `mutate()` to add standardized scores

```{r, eval = FALSE}
Schooling.mean = mean(life1$Schooling, na.rm = TRUE)
Schooling.sd = sd(life1$Schooling, na.rm = TRUE)

life1 <- life1 %>% 
  mutate(
      Schooling.std = (Schooling - Schooling.mean)/Schooling.sd
      )

ggplot(life1, aes(x = Schooling.std))+
    geom_density()+
    theme_bw()
```

Can you add another column to include the standardized scores for BMI? 

---

### Task 5: Examine one case

Let's examine one particular country, Australia, for the year 2010. Interpret the z-scores of Schooling and say how it compares in light of the distribution of the two variables.

**HINT: Use the `filter()` function**

General syntax - 

```{r, eval = FALSE}
new_df <- filter(
  old_df, #name of dataset
  some condition on the columns 
  #names of columns as it appears and without quotes
)

#Or,
new_df <- old_df %>% 
    filter(some condition on the columns)
```

---

### One way to do Task 5

```{r, eval = FALSE}
life1_aus <- life1 %>% 
    filter(Country == "Australia" & Year == "2011")

head(life1_aus)

```