---
title: "Week 2 Lab"
author: "(add your name)"
date: "1/11/2023"
output: 
 html_document:
     code_folding: hide
---

```{r setup, include=FALSE}
# Don't make a change here
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


### Task 1: Load the libraries and read the data

```{r}
# For this lab, you will need the following libraries: 
# haven, tidyverse, modelsummary
# load them in your environment by completing the blanks
# Remember to uncomment (remove the hash) the code for it ru

# library(.......)
# ....
# ....
```


```{r}
# Saving the data in an eat object

# eat <- read_spss("https://daviddliebowitz.github.io/EDUC643_23W/data/male_do_eating.sav") %>%
#     select(OE_frequency, EDEQ_restraint, EDS_total, BMI, age_year, income_group)

# Print the head of the data
# Print the str of the data

# head(...)
# ....
```

**Q1. Describe the structure of the data in a few lines.**

*<Enter answer here>*

### Task 2 - Descriptive statistics

```{r}
# cleaning the data

# eat <- eat %>%
#     filter(!is.na(OE_frequency), !is.na(EDS_total)) %>%
#     mutate(EDS_total = ifelse(EDS_total==-99, NA, EDS_total)) %>%
#     drop_na()

# Find some descriptive statistics for the variables - BMI and EDEQ_restraint

# .....
# .....
    
```

**Q2. Report the descriptive statistics for BMI and EDEQ_restraint.**

*<Enter answer here>*

### Task 3 - Bi-variate Relationship

We are interested in how dietary restraint (EDEQ_restraint) predicts a person's body-mass index (BMI). 

**Q3. What are the dependent and independent variables here?**

*<Enter answer here>*

```{r}
# Make a bivariate plot. Fill what should be on the x and y-axis.
# And give a meaningful title

# lm_plot <- ggplot(data=do, aes(x= ....., y= .....)) + 
#               geom_point() +
#               labs(
#               x = "Dietary restraint index (0-6)",
#               title = "") +
#               theme_minimal(base_size = 16) +
#               geom_smooth(method='lm', se=F)

# lm_plot
```

**Q4. Describe the bi-variate relationship using direction, linearity, outliers, strength, and magnitude.**

*<Enter answer here>*

```{r}
# Run an OLS linear regression model
# Write the function, and the formula

#eat_model <- ...(... ~ ..., data = eat)

#summary(eat_model)

```

**Q5. Describe the model result. Report and interpret the intercept and slope. Comment if the relationship is statistically significant**

### Task 4 - Summarize

**In this lab, we explored the relationship between dietary restraint and BMI. To do so, we calculated summary statistics, plotted the bi-variate relationship, and ran an OLS linear regression model. To explain your work to an audience, write a paragraph synthesizing all the pertinent informaiton.**

*<Enter answer here>*

### Challenge

**Choose two continuous variables from the data set that you find interesting. Explore and report their bi-variate relationship.**

