---
title: "Week 3 Lab"
author: ""
date: "1/23/2023"
output: 
 word_document
---

Hello, welcome to EDUC 643 week 3. To allocate more time for you to work on assignment 1 and obtain individual assistance from your GEs, this worksheet is finished in terms of coding (text answers to substantive questions are left out; you're more than welcome to discuss these with your classmates or GEs during the lab).

To use this worksheet, follow the steps:

  - Knit this Rmd file first and check out the output. You should see a word document that contains the output **elements** that are expected in your assignment 1 submission (only the dataset here is the BMI data)
  - Go back to this Rmd file, find the code that you need for some specific **elements**, and rewrite the code in a way that works for your assignment 1
  - If you have any issues using the code in this Rmd file, ask your GEs for assistance
  

```{r setup, include=FALSE}
# Don't make a change here
knitr::opts_chunk$set(include = TRUE,
                      echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

### Task 1: Load the libraries and read the data

For this lab, you will need the following libraries: haven, tidyverse, modelsummary, and flextable. 

What's new today about these packages:

 - We'll use the modelplot() function of the {modelsummary} package to generate nice plot to visualize coefficients and corresponding confidence intervals in one code line. 
 - We'll use the "output = flextable" argument in the modelsummary() function to create more, I mean, the package {flextable} is kind of self-explaining: more flexible tables.

```{r}
library(haven)
library(tidyverse)
library(modelsummary)
library(flextable)
```


```{r, include = FALSE}
# Saving the data in an eat object

eat <- read_spss("https://daviddliebowitz.github.io/EDUC643_23W/data/male_do_eating.sav") %>%
    select(OE_frequency, EDEQ_restraint, EDS_total, BMI, age_year, income_group)

# Print the head of the data
# Print the str of the data

head(eat)
str(eat)
```

Q1. Describe the structure of the data in a few lines: How many observations? How many variables? For each variable, is it currently continuous or categorical? Should you recode it?

### Task 2 - Descriptive statistics

```{r}
# cleaning the data
eat <- eat %>% 
  mutate(EDS_total = ifelse(EDS_total==-99, NA, EDS_total)) %>%
  drop_na()

# Find some descriptive statistics for the variables - BMI and EDEQ_restraint

eat %>% 
  select(BMI, EDEQ_restraint) %>% 
  datasummary_skim()
```

Q2. Report the descriptive statistics for BMI and EDEQ_restraint: since they are all continuous variables, report and interpret their mean and standard deviation for the minimum, and report other statistics as you see fit.

### Task 3 - Bi-variate Relationship

We are interested in how dietary restraint (EDEQ_restraint) predicts a person's body-mass index (BMI). 

Q3. What are the dependent and independent variables here?

```{r}
# Make a bivariate plot. Fill what should be on the x and y-axis.
# And give a meaningful title

lm_plot <- ggplot(data=eat, aes(x= EDEQ_restraint, y= BMI)) +
  geom_point() +
  labs(x = "Dietary restraint index (0-6)",
       y = "BMI",
       title = "Bivariate relationship between dietary restraint and BMI") +
  theme_minimal(base_size = 16) +
  geom_smooth(method='lm', se=F)

lm_plot
```

Q4. Based on the plot, describe the bi-variate relationship using direction, linearity, outliers, strength, and magnitude.


Q5. Write a formal linear model that describes the simple, bivariate relationship

$$ BMI = \beta_0 + \beta_1 * EDEQ_restraint + \epsilon $$
Q6. Fit a linear model using OLS approach and report the results

**Important note**: the table below reports $R^2$, $RMSE$, $F$ statistic, and confidence intervals only because you're now learning these concepts, not because they are common practice among researchers. 

```{r}
# Run an OLS linear regression model
# Write the function, and the formula

m1 <- lm(BMI ~ EDEQ_restraint, data = eat)

modelsummary(m1,
             output = 'flextable',
             stars = TRUE,
             estimate  = c("{estimate} {stars} [{conf.low}, {conf.high}]"),
             gof_omit = "Adj.|Log.Lik.|AIC|BIC",
             coef_rename = c('EDEQ_restraint' = 'Dietary restraint')) %>% 
  autofit()

```

Q7. (This is completely extra and not required in the assignment) Visualize confidence interval

```{r}
lm_plot +
  geom_smooth(method='lm', fill = "red")
```


```{r}
modelplot(m1)
```
