---
title: "Week 8 Lab (Answer Sheet)"
author: "(enter your name)"
date: "2/27 or 2/28"
output: 
    html_document:
        code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{css, echo=FALSE}
/* Whole document: */
body{
  font-family: Helvetica;
  font-size: 13pt;
}
```


```{r load in packages}
pacman::p_load(tidyverse, modelsummary, margins)
```

## Read in DIBELS dataset

```{r read in data}
df0 <- read.csv("https://daviddliebowitz.github.io/EDUC643_23W/data/dibels.csv") 
```

```{r}
df1 <- df0 %>% 
  select(id = sch_deid,
         grade,
         w1 = y1_boy_mean,
         w2 = y1_moy_mean,
         w3 = y2_boy_mean,
         w4 = y2_moy_mean,
         enrollment = school_enroll,
         frpl = frpl_prop)

df <- pivot_longer(df1,
                   cols = w1:w4,
                   names_to = "wave",
                   names_prefix = "w",
                   values_to = "score") %>% 
  mutate(id = as.character(id),
         post = as.factor(ifelse(wave %in% c(1, 2), 0, 1)),
         grade = as.factor(grade),
         enrollment = as.numeric(enrollment))
```

In this lab, we continue using [the DIBELS dataset](https://daviddliebowitz.github.io/EDUC643_23W/data/dibels.csv) from lab 7 to check our understanding of multiple regression (ML), where outcome variable is continuous and predictor of research interest is a continuous variable AND its interaction term with another continuous/categorical variable.

## Intuition

We start with our old friend model 1, where our research inquiry is the relationship between outcome Y and predictor X, and our coefficient of interest is $\beta_1$ - the coefficient on X. 

Now we go a step further: we theorize that another variable Z will change the relationship between Y and X, so we fit our data to model 2 and we are now interested in both $\beta_1$ and $\beta_3$ - the slope and the change in slope brought upon by Z. Additionally, we may be also interested in $\beta_0$ and $\beta_2$ - the intercept and the change in intercept brought upon by Z.

Model 1:
$$ Y = \beta_0 + \beta_1X + \epsilon $$

```{r, echo=FALSE}
ggplot(df, aes(score, frpl)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "X",
       y = "Y") +
  theme_bw(base_size=16) 
```


Model 2:
$$ Y = \beta_0 + \beta_1X + \beta_2Z + \beta_3X:Z + \epsilon $$

```{r, echo=FALSE}
df %>% 
  mutate(Z = as.factor(ifelse(wave == 1 | wave == 3, 1, 0))) %>% 
  ggplot(aes(score, frpl, color = Z)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "X",
       y = "Y") +
  theme_bw(base_size=16) 
```

Hopefully, after this lab, you'll figure out how to use OLS regression to help you find out the intercept and slope for the best fit line of Z = 0 as well as the intercept and slope for the best fit line of Z = 1. Imagine Z = 0 vs Z = 1 is the condition of control vs treatment, how helpful this skill will be to your understanding of the effect of an intervention. 

Putting the models and plots together, for your curiosity, 

 - the intercept and slope for line Z = 0 are $\beta_0$ and $\beta_1$
 - the intercept and slope for line Z = 1 are $\beta_0 + \beta_2$ and $\beta_1 + \beta_3$
 - the change in intercept is $\beta_2$ and the change in slope is $\beta_3$
 
Making sense? Let's dive into research questions and analyses.

## Key variables

We conducted some data management on the original DIBELS data and currently we have these following variables:

our outcome variable Y (continous, required by OLS)

 - score: ORF score (school-by-grade-by-wave level)

our predictor variable X (continous for this lab, check the lecture code for categorical predictor)

 - frpl: proportion (school-by-grade level) of students receiving free or reduced price school lunch 

our theoretical third variable Z (categorical or continous) that may change the Y-X relationship 

 - post: coded one for school year 2020-21 and zero for school year 2019-20
 - grade: coded 1-5 for 1-5 graders
 - enrollment: test enrollment (school level) 

Descriptives of continuous variables:

```{r}
df <- df %>% 
  select(-wave)

datasummary_skim(df)
```

Descriptives of categorical variables:

```{r}
datasummary_skim(df, type = "categorical")
```

## Research questions

Q1. How did students' Oral Reading Fluency (ORF) trajectories differ pre- and post-pandemic-onset?

Q2. To what extent did differences in students' ORF trajectories pre- and post-pandemic-onset differ by the proportion of students receiving free- or reduced-price lunch (FRPL) in their schools?

Q3. To what extent did differences in students' ORF trajectories pre- and post-pandemic-onset differ by grade level?

Q4. To what extent do students' average ORF scores differ by the rate of FRPL-receipt and enrollment in their schools?

## Data analysis

We use OLS regression to estimate four linear models in the following.


```{r}
fit1 <- lm(score ~ post, df)
fit2 <- lm(score ~ post*frpl, df)
fit3 <- lm(score ~ post*grade, df)
fit4 <- lm(score ~ frpl*enrollment, df)
```


Q1 How did students' Oral Reading Fluency (ORF) trajectories differ pre- and post-pandemic-onset?

$$SCORE = \beta_0 + \beta_1POST + \epsilon$$

```{r}
modelsummary(fit1,
             stars = TRUE,
             gof_omit = "Adj.|Log.Lik.|AIC|BIC")
```

Compared to pre-pandemic, the average score was ____ (higher/lower) (p_____) post pandemic.


Q2. To what extent did differences in students' ORF trajectories pre- and post-pandemic-onset differ by the proportion of students receiving free- or reduced-price lunch (FRPL) in their schools?

$$SCORE = \beta_0 + \beta_1POST + \beta_2FRPL + \beta_3POST:FRPL + \epsilon$$


```{r}
modelsummary(fit2,
             stars = TRUE,
             gof_omit = "Adj.|Log.Lik.|AIC|BIC")
```


In schools where no student receives free/reduced lunch (frpl = 0 or 0%), 

 - intercept: ____, is the pre-pandemic average score
 - slope: ____, is the change in score after pandemic, compared to pre-pandemic

Careful - in schools where all students receives free/reduced lunch (frpl = 1 or 100%) (hint: plug frpl = 1 into your model), 

 - intercept: ____, is the pre-pandemic average score
 - slope: ____, is the change in score after pandemic, compared to pre-pandemic
 
Taking together, what is the change in slope (i.e., change in score from pre to post-pandemic) that is related to one unit change in the proportion of student receiving free/reduced lunch? is it statistically significant? What is your answer to Q2?
 

Q3. To what extent did differences in students' ORF trajectories pre- and post-pandemic-onset differ by grade level?

$$SCORE = \beta_0 + \beta_1POST + \beta_2GRADE + \beta_3POST:GRADE + \epsilon$$

```{r}
modelsummary(fit3,
             stars = TRUE,
             gof_omit = "Adj.|Log.Lik.|AIC|BIC")
```


For grade 1 students, 

 - intercept: ____, is the pre-pandemic average score
 - slope: ____, is the change in score after pandemic, compared to pre-pandemic

Careful - for grade 2 students (hint: plug these into your model, grade2 = 1, grade3 = 0, grade4 = 0, grade5 = 0), 

 - intercept: ____, is the pre-pandemic average score
 - slope: ____, is the change in score after pandemic, compared to pre-pandemic

Careful - for grade 3 students, 

 - intercept: ____, is the pre-pandemic average score
 - slope: ____, is the change in score after pandemic, compared to pre-pandemic

Careful - for grade 4 students, 

 - intercept: ____, is the pre-pandemic average score
 - slope: ____, is the change in score after pandemic, compared to pre-pandemic

Careful - for grade 5 students, 

 - intercept: ____, is the pre-pandemic average score
 - slope: ____, is the change in score after pandemic, compared to pre-pandemic

Taking together, what is the difference in slope (i.e., difference in score from pre to post-pandemic) between grade 1 and grade 2? what is the substantive meaning of this number? is it statistically significant? is it one of the answers to Q3? What is your full answer to Q3?

Q4. To what extent do students' average ORF scores differ by the rate of FRPL-receipt and enrollment in their schools?

$$SCORE = \beta_0 + \beta_1FRPL + \beta_2ENROLLMENT + \beta_3FRPL:ENROLLMENT + \epsilon$$
```{r}
modelsummary(fit4,
             stars = TRUE,
             gof_omit = "Adj.|Log.Lik.|AIC|BIC")
```

In schools where no student receives free/reduced lunch (frpl = 0 or 0%), 

 - intercept: ____, is the average score for schools with no enrollment
 - slope: ____, is the change in score related to one student enrollment at school 

Careful - in schools where all students receives free/reduced lunch (frpl = 1 or 100%), 

 - intercept: ____, is the average score for schools with no enrollment
 - slope: ____, is the change in score related to one student enrollment at school
 
Taking together, what is the difference in slope (i.e., difference in score from pre to post-pandemic) for schools with frpl = 0 vs schools with frpl = 1? what is the substantive meaning of this number? is it statistically significant? is it the answer to Q4?

